---
title: "Hydrological Modeling in Atbashy River Basin: Hydrological Modeling (airGR)"
author: "Tobias Siegfried, hydrosolutions GmbH"
format: html
editor: source
date: "2025-04-30"
---

# I. BACKGROUND

## Hydrological Modeling in Atbashy River Basin

### Introduction

This script implements a hydrological modeling framework to simulate discharge in Atbashy River Basin, Kyrgyzstan, using the airGR package. The modeling framework is designed to integrate various data sources and methodologies to provide a comprehensive understanding of the hydrological processes in the basin. The primary focus is on simulating the hydrological response to 2017 - 2023 forcing, i.e., the years of the focus of the study by Hannah Schwedhelm.

#### Key Components

1. **Model Structure**
   - Uses the airGR package's CemaNeige model coupled with GR4J
   - Implements elevation band discretization to account for orographic effects
   - Incorporates glacier melt contributions using Rounce et al. (2023) projections
   - Handles snow/rain partitioning using the Glazirin method calibrated for Central Asian conditions

2. **Data Integration**
   - Historical climate observations for model forcing
   - Monthly discharge measurements for calibration and validation of mean monthly flows
   - Glacier melt data with temperature-based temporal disaggregation
   - Elevation band-specific precipitation and temperature data

3. **Modeling Framework**
   - Split-sample calibration and validation approach
   - Multi-criteria performance evaluation
   - Latin Hypercube Sampling for parameter uncertainty analysis
   - Specific focus on snow/rain transition parameterization

#### Objectives

The primary objectives of this modeling exercise are to:
1. Establish a robust hydrological model for the Atbashy River Basin
2. Account for the significant role of snow and glacier melt in the basin's hydrology
3. Provide insights into parameter sensitivity and model uncertainty

This script is a spin off of a larger climate change impact assessment project and builds upon previous data preparation work documented in separate scripts. It is the first time that we try to implement a full modeling chain in R only without RSMINERVE but with airGR. The script is designed to be modular and adaptable, allowing for easy updates and modifications as new data becomes available or as modeling techniques evolve.

Challenge: We only have monthly observations and airGR presumes daily observations.How to deal with this?


# II. Change Log
- **2024-11-19**:
  - Initial setup.
- **2024-11-20**:
  - Integration with the data preparation script.
- **2024-11-21**:
  - Simple monthly to daily glacier melt data disaggregation implemented.
  - Testing the teaching package.
- **2024-11-24**:
  - continuing investigation into CEMA-Neige.
- **2024-11-25**: 
  - Way forward is clear. airGR::CreateInputsModel produces elevation bands specific model input. We can just overwrite the results from the data_preparation.qmd script where we have prepared elevation band data for P and T. 
- **2024-11-26**:
  - Continuing with hydro modeling using airGR.
  - The temperature to peak discharge sensitivity is stunning. We get a good calibration during the summer months if we lower the 'observed' temperatures by 10°C. On the plot with the rolling mean, the calibrated peak discharge coincides well with the observed discharge. However, under such scheme, the winter discharge is mostly overestimated. This sensitivity shows the importance of 'bias correcting' the CHELSA V21 temperature fields.
- **2024-11-28**:
  - Progressing on climate impact implementation
  - functions now in separate R scripts. The deep rationale for this was to be able to provide them as context to Claude. It does not work with .qmd script files.
- **2024-11-29**:
  - Running climate simulations now and starting to analyze outputs.
  - Export of relevant results for later analysis.    
- **2024-12-02**:
  - Streamlining code and improving documentation.
- **2024-12-03**:
  - Need for resetting snow every 5 years to avoid issue with snow accumulation in high elevation catchments. 
- **2024-12-04**:
  - Code updated. Results looking good.
  - Added shiny figure plotting.
- **2024-12-06**:
  - Finalizing the script for error checking. It fully replaces the climate impact analysis script.
- **2025-02-27**:
  - Validating code and crosschecking if all works.
- **2025-03-06**:
  - Preparing script for use with Positron IDE.
- **2025-04-28**:
  - Adaptation for Atbashy hydrological modeling study
- **2025-04-29**:
  - Working on the calibration of the GR4J model with daily inputs and monthly discharge observations.
- **2025-04-30**:
  - Switched to LHS optimization for parameter calibration. First Claude Version available for testing now.

# ===============================

# 1. CONFIGURATIONS & FUNCTIONS

## 1.0 Configurations

```{r}
# Load the centralized configuration
library(pacman)
p_load(here, tidyverse, lubridate)

path = here::here("01_code", "config_param.R")
source(path)
```

## 1.1 Source functions

```{r}
source(here::here("01_code", "functions", "disaggregate_glacier_melt.R"))
source(here::here("01_code", "functions", "solid_fraction_elevation_layer.R"))
source(here::here("01_code", "functions", "process_rsminerve_climate_files.R"))
source(here::here("01_code", "functions", "process_glacier_data.R"))
source(here::here("01_code", "functions", "plot_calib_valid_period.R"))
source(here::here("01_code", "functions", "create_basin_scenario_model_ts.R"))
source(here::here("01_code", "functions", "process_inputs_models_scenarios.R"))
source(here::here("01_code", "functions", "run_climate_scenarios.R"))
source(here::here("01_code", "functions", "plot_flexible_hydrograph.R"))
source(here::here("01_code", "functions", "calculate_exceedance.R"))
source(here::here("01_code", "functions", "plot_exceedance.R"))
source(here::here("01_code", "functions", "adaptive_lhs_calibration.R"))
source(here::here("01_code", "functions", "parameter_space_visualization_functions.R"))
```

# 2. PARAMETERS & DATA

## 2.1 Parameters

Loading data that was produced in the data preparation script. This includes the basin information and the calibration/validation periods. 

```{r}
# Basin information
Basin_Info <- read_rds(file.path(config$paths$data_path, "Basin_Info.rds"))
Basin_Info |> summary()

# time periods and calibration/validation periods
cal_val_per <- readRDS(file.path(config$paths$data_path, "Calibration_Validation_Period.rds"))
cal_val_per |> summary()

# Snow reset level and threshold. This is the level at which the snow is reset in the model at the end of a simulation period chunk. This is important to avoid consistent snow accumulation in high elevation catchments. All is in mm.
snow_reset_level <- 0 # mm
snow_reset_threshold <- 1000 # mm
```

## 2.2 Data

### Discharge Data (HIST_OBS)

Setting up monthly discharge data. 

```{r}
#| warning: false

q <- readRDS(file.path(paste0(config$paths$data_path,"/Discharge"), "q_cal_val.rds"))

# set date column to first day of month
q <- q |> 
  mutate(date = ymd(date)) |> 
  mutate(date = floor_date(date, "month"))

# Add specific discharge in mm/day
q <- q |> 
  rename(Q_m3_sec = value) |>
  mutate(Q_mm_day = Q_m3_sec / as.numeric(Basin_Info$BasinArea_m2) * 10^3 * 3600 * 24) |> 
  dplyr::select(-data)

# Add/Update period information
## Filter out data prior to 1979-01-01
q <- q[complete.cases(q), ]

## Set warmup period
n_warmup_years <- 3

## reset labels
q <- q %>%
  mutate(Period = case_when(
    date < cal_val_per$calib_start + years(n_warmup_years) - days(1) ~ "Warmup",
    date >= cal_val_per$calib_start & date <= cal_val_per$calib_end ~ "Calibration",
    date > cal_val_per$valid_start & date <= cal_val_per$valid_end ~ "Validation"
  ))

plot_calib_valid_period(q)

# summarize by time compute annual discharge stats
q |> 
  group_by(Period) |>  
  timetk::summarize_by_time(.date_var = date, .by = "year", 
                            Q_mm_day = sum(Q_mm_day) * 3600 * 24 * 365.25 / 10^9) |> 
  summary()
```

### Forcing Data (HIST_OBS)

Here, we load hist_obs forcing data and then average over the elevation bands. 
Depending on the availability of station data, we can correct the CHELSA V21 forcing fields. 

```{r}
#| warning: false
#| echo: false

hist_obs <- read_csv(file.path(config$paths$model_dir, "forcing/hist_obs_rsm.csv"), 
                     col_names = FALSE,
                     show_col_types = FALSE) 
# delete the last column in hist_obs (discharge measured at the gauge) for hist_obs
hist_obs <- hist_obs %>% dplyr::select(-ncol(hist_obs)) 

hist_obs_processed <- process_rsminerve_climate_files(hist_obs)

dates_tbl <- hist_obs_processed$dates
hist_obs_T_bands <- hist_obs_processed$T_bands
hist_obs_P_bands <- hist_obs_processed$P_bands

hist_obs_T <- hist_obs_T_bands |> rowMeans()
hist_obs_P <- hist_obs_P_bands |> rowMeans()
```

### Glacier Data (HIST_OBS)

Subtract the monthly Rounce et al. 2023 data from the monthly discharge data of the hist_obs period and create net discharge time series.

```{r}
#| warning: false
#| message: false
#| echo: false

# Load the glacier data
qg_mon <- read_csv(file = file.path(config$paths$glacier_path, "glaciers_hist_obs_rsm.csv"))
qg_mon <- qg_mon |> 
  slice(-1:-7)# Remove the first 7 rows

# Now rename and calculate
qg_mon <- qg_mon %>% 
  rename(Q_m3_sec = `Glacier 16076`) %>%  # Note the backticks for column names with spaces
  rename(date = Station) |> 
  mutate(Q_m3_sec = as.numeric(Q_m3_sec),  # Ensure the column is numeric
         Q_mm_day = Q_m3_sec / as.numeric(Basin_Info$BasinArea_m2) * 10^3 * 3600 * 24)

monthly_melt <- qg_mon |> 
  dplyr::select(date, Q_mm_day) |> rename(melt = Q_mm_day) |> 
  mutate(date = dmy_hms(date)) |> 
  mutate(date = date(date)) #|>
  #mutate(date = ceiling_date(date, "month") - days(1))

```

### Discharge Net Glacier Melt

```{r}
q <- q |> 
  left_join(monthly_melt, by = "date") |> 
  mutate(melt = ifelse(is.na(melt), 0, melt)) |> 
  mutate(Q_net_melt_mm_day = Q_mm_day - melt)

# plot Q_net_melt_mm_day time series
q |> 
  dplyr::select(date, Q_net_melt_mm_day, melt) |> 
  pivot_longer(-date) |> 
  #group_by(name) |> 
  timetk::plot_time_series(.date_var = date, .value = value, 
                   .smooth = FALSE, .color_var = name,
                   .title = "Discharge without Glacier Contribution")
```

### Final Dataframe via combination of data

```{r}
#| warning: false
#| message: false
#| echo: false

# Combine the data
forcing_q_ts <- data.frame(date = dates_tbl$value, 
                         #julian day
                         JourJul = yday(dates_tbl$value),
                         Ptot = hist_obs_P, 
                         Temp = hist_obs_T) 

# add discharge

q_net_melt_mm_day <- q %>% dplyr::select(date,Q_net_melt_mm_day)

forcing_q_ts <- forcing_q_ts |> 
  left_join(q_net_melt_mm_day, by = "date") |> 
  rename(Q_mm_day = Q_net_melt_mm_day)

# compute potential ET using Oudin method
PET = airGR::PE_Oudin(JD = forcing_q_ts$JourJul, 
                      Temp = forcing_q_ts$Temp, 
                      Lat = Basin_Info$BasinLat_rad, 
                      LatUnit = "rad", 
                      TimeStepIn = "daily", 
                      TimeStepOut = "daily")

# bind all together
forcing_q_ts <- cbind(forcing_q_ts, PET = PET)

# visualize
forcing_q_ts |> 
  dplyr::select(date, Q_mm_day, Ptot, Temp, PET) |>
  pivot_longer(cols = c(Q_mm_day, Ptot, Temp, PET), names_to = "Variable", values_to = "Value") |> 
  # generate an interactive plot
  ggplot(aes(x = date, y = Value, group = Variable)) +
  geom_line(aes(color = Variable)) +
  geom_point(aes(color = Variable)) +
  labs(x = "Date", y = "Value", color = "Variables") +
  theme_minimal()

# save as csv on disc
write_csv(forcing_q_ts, file.path(config$paths$model_dir, "forcing_q_ts.csv"))
```

# 3. MODEL

## 3.0 Model Specific Setup
```{r}
# Load the data
basin_obs_ts <- forcing_q_ts |> 
  rename(Q_mm_day_no_glacier = Q_mm_day)

# setting calib_end
cal_val_per$calib_end <- ymd(cal_val_per$calib_end)

# indexing calibration time steps
indRun_cal <- which(basin_obs_ts$date >= cal_val_per$calib_start & 
                      basin_obs_ts$date <= cal_val_per$calib_end)
indRun_cal_max <- indRun_cal |> max()

# model to use 
model_2_use <- "GR4J"

if (model_2_use == "GR4J") {
  fun_model <- airGR::RunModel_CemaNeigeGR4J
} else if (model_2_use == "GR5J") {
  fun_model <- airGR::RunModel_CemaNeigeGR5J
} else if (model_2_use == "GR6J") {
  fun_model <- airGR::RunModel_CemaNeigeGR6J
}

# warmup period
n_warmup_years <- 3
warmup_period_ind <- 1:(365*n_warmup_years + 1)
# calibration period
indRun_cal <- indRun_cal + length(warmup_period_ind)
indRun_cal <- indRun_cal[indRun_cal<=indRun_cal_max] # this makes sure that we remain < calib_end period
# validation period
indRun_val <- which(basin_obs_ts$date >= cal_val_per$valid_start & 
                      basin_obs_ts$date <= cal_val_per$valid_end)

# transformation
transfo = ""

# performance criteria
fun_crit = airGR::ErrorCrit_NSE
weights = c(1)

#fun_crit = list(ErrorCrit_NSE, ErrorCrit_RMSE)
#weights = c(1/2, 1/2)

# testing custom function
#fun_crit = ErrorCrit_NSE_Monthly
#weights = c(1)

# observation
q_obs_cal = basin_obs_ts$Q_mm_day_no_glacier[indRun_cal]
q_obs_val = basin_obs_ts$Q_mm_day_no_glacier[indRun_val]
```


## 3.1 Prepare Model

Note: We ran the parameter optimization experiment in the X.EXPERIMENT section below. The best parameters for tas_min and tas_max were found to be 0.213 and 0.489, respectively.

```{r}
basin_obs_ts <- forcing_q_ts

# preparation of input data
inputsModel <- airGR::CreateInputsModel(FUN_MOD            = fun_model, 
                                        DatesR             = basin_obs_ts$date,
                                        Precip             = basin_obs_ts$Ptot, 
                                        PotEvap            = basin_obs_ts$PET,
                                        TempMean           = basin_obs_ts$Temp, 
                                        HypsoData          = Basin_Info$HypsoData,
                                        ZInputs            = median(Basin_Info$HypsoData),
                                        NLayers            = length(Basin_Info$ZLayers))

# overwrite airGR CemaNeige elevation bands data with our own.
inputsModel_ours <- inputsModel
inputsModel_ours$LayerTempMean <- hist_obs_T_bands |> as.list() 
inputsModel_ours$LayerPrecip <- hist_obs_P_bands |> as.list()

# compute the fraction of solid precipitation on the layer. # tas_min and tas_max are the best parameters found in the experiment under Section X below.

tas <- c(0.213, 0.489) # tas_min and tas_max are the best parameters found in the experiment under Section X below.
#tas <- c(9.1,10)

solid_frac_pr <- solid_fraction_elevation_layer(Basin_Info,
                                                  inputsModel_ours$LayerTempMean, 
                                                  tas_min  = tas[1], 
                                                  tas_max  = tas[2]) 

inputsModel_ours$LayerFracSolidPrecip <- solid_frac_pr
# overwrite original
inputsModel <- inputsModel_ours

# What we are doing here is to overwrite the setup prepared in the airGR::CreateInputsModel() script. In other words, we just use the data from the data_preparation script.
runOptions_cal <- airGR::CreateRunOptions(FUN_MOD          = fun_model,
                                          InputsModel      = inputsModel,
                                          IndPeriod_Run    = indRun_cal,
                                          IniStates        = NULL,
                                          IniResLevels     = NULL,
                                          IndPeriod_WarmUp = warmup_period_ind,
                                          IsHyst           = FALSE,
                                          warnings         = FALSE,
                                          Outputs_Sim      = "Qsim")
 
inputsCrit_cal <- airGR::CreateInputsCrit(FUN_CRIT         = fun_crit,
                                          InputsModel      = inputsModel,
                                          RunOptions       = runOptions_cal,
                                          transfo          = transfo,
                                          Obs              = q_obs_cal,
                                          warnings         = FALSE,
                                          Weights          = weights)

calibOptions <- airGR::CreateCalibOptions(FUN_MOD          = fun_model, 
                                          FUN_CALIB        = airGR::Calibration_Michel)
```


# 4. MODEL CALIBRATION

## Step 1: Load Required Libraries

Load data and prepare model structure (see above).

## Step 2: Prepare Monthly Observations

```{r}
# Load discharge data
#q <- readRDS(file.path(paste0(config$paths$data_path,"/Discharge"), "q_cal_val.rds"))
#q <- q |> 
#  mutate(date = ymd(date)) |> 
#  mutate(date = floor_date(date, "month")) |>
#  rename(Q_m3_sec = value) |>
#  mutate(Q_mm_day = Q_m3_sec / as.numeric(Basin_Info$BasinArea_m2) * 10^3 * 3600 * 24)

# Extract calibration period observations
obs_monthly_cal <- q %>%
  filter(date >= cal_val_per$calib_start & date <= cal_val_per$calib_end)
```

## Step 3: Define Parameter Ranges

```{r}
# For GR4J + CemaNeige
param_ranges <- list(
  X1 = c(100, 12000),            # Production store capacity (mm)
  X2 = c(-5, 5),                # Groundwater exchange coefficient (mm/day)
  X3 = c(20, 300),              # Routing store capacity (mm)
  X4 = c(1, 10),               # Unit hydrograph time constant (days)
  CemaNeigeParam1 = c(0.5, 2),  # Snow melt temperature factor
  CemaNeigeParam2 = c(0.5, 200)   # Snow cold content factor
)
```

## Step 4: Create Model Run and Evaluation Functions

```{r}
# Function to run the model with parameters
run_model_wrapper <- function(params) {
  # Convert parameters list to vector
  param_vector <- unlist(params)
  
  # Run model
  results <- airGR::RunModel_CemaNeigeGR4J(
    InputsModel = inputsModel,
    RunOptions = runOptions_cal,
    Param = param_vector
  )
  
  return(results)
}

# Function to evaluate monthly performance
eval_function_wrapper <- function(model_results) {
  # Extract simulated discharge
  Qsim <- model_results$Qsim
  
  # Aggregate to monthly
  sim_df <- data.frame(
    date = basin_obs_ts$date[indRun_cal],
    Qsim = Qsim
  )
  sim_df$month_key <- format(sim_df$date, "%Y-%m")
  sim_monthly <- aggregate(Qsim ~ month_key, data = sim_df, FUN = mean, na.rm = TRUE)
  
  # Format observations
  obs_df <- data.frame(
    month_key = format(obs_monthly_cal$date, "%Y-%m"),
    Qobs = obs_monthly_cal$Q_mm_day
  )
  
  # Merge and calculate NSE
  merged_data <- merge(sim_monthly, obs_df, by = "month_key")
  
  # Calculate NSE
  mean_obs <- mean(merged_data$Qobs, na.rm = TRUE)
  numerator <- sum((merged_data$Qsim - merged_data$Qobs)^2, na.rm = TRUE)
  denominator <- sum((merged_data$Qobs - mean_obs)^2, na.rm = TRUE)
  
  if (denominator > 0) {
    nse <- 1 - numerator / denominator
  } else {
    nse <- -999
  }
  
  return(nse)
}
```

## Step 5: Run Adaptive Calibration with LHS

```{r}
# Run calibration

set.seed(111)

calibration_results <- adaptive_lhs_calibration(
  run_model = run_model_wrapper,
  eval_function = eval_function_wrapper,
  param_ranges = param_ranges,
  n_initial_samples = 1000,    # Number of initial samples
  n_iterations = 100,          # Number of refinement iterations
  refinement_factor = .1,   # Top 10% used for refinement
  n_refined_samples = 100,    # Samples per refinement
  parallel = FALSE,           # Set to TRUE for parallel processing
  n_cores = 9
)

# Extract best parameters
best_params <- calibration_results$best_params
best_performance <- calibration_results$best_performance

# Display results
print(best_params)
print(paste("Best NSE:", best_performance))
```
Run model with best parameters.

```{r}
# After extracting best parameters from calibration_results
best_params <- calibration_results$best_params
param_vector <- unlist(best_params)  |> unname()# Convert from list to vector

# new run options to output everything
runOptions_run <- airGR::CreateRunOptions(FUN_MOD          = fun_model,
                                          InputsModel      = inputsModel,
                                          IndPeriod_Run    = indRun_cal,
                                          IniStates        = NULL,
                                          IniResLevels     = NULL,
                                          IndPeriod_WarmUp = warmup_period_ind,
                                          IsHyst           = FALSE,
                                          warnings         = FALSE,
                                          Outputs_Sim      = "all")

# Choose the appropriate model function based on your model_type
if (model_2_use == "GR4J") {
  best_run_cal <- airGR::RunModel_CemaNeigeGR4J(
    InputsModel = inputsModel,
    RunOptions = runOptions_run,
    Param = param_vector
  )
} else if (model_2_use == "GR5J") {
  best_run_cal <- airGR::RunModel_CemaNeigeGR5J(
    InputsModel = inputsModel,
    RunOptions = runOptions_run,
    Param = param_vector
  )
} else if (model_2_use == "GR6J") {
  best_run_cal <- airGR::RunModel_CemaNeigeGR6J(
    InputsModel = inputsModel,
    RunOptions = runOptions_run,
    Param = param_vector
  )
}
```

## Step 6: Visualize Calibration Results

```{r}
#| warning: false
#| message: false
#| echo: false

# Plot parameter sensitivity
plot_parameter_sensitivity(calibration_results$param_df)

# Plot parameter correlations
plot_parameter_correlations(calibration_results$param_df)

# Plot parameter evolution
plot_parameter_evolution(calibration_results$param_df)

# Plot monthly comparison
plot_monthly_comparison(
  best_run_cal,
  obs_monthly_cal,
  basin_obs_ts$date[indRun_cal],
  "Calibration Period"
)
```

## Step 7: Validate the Model

```{r}
#| warning: false
#| message: false
#| echo: false

# Run validation
runOptions_val <- airGR::CreateRunOptions(
  FUN_MOD = fun_model,
  InputsModel = inputsModel,
  IndPeriod_Run = indRun_val,
  IniStates = best_run_cal$StateEnd,
  IniResLevels = NULL
)

# Run model with best parameters
param_vector <- unlist(best_params)
best_run_val <- airGR::RunModel_CemaNeigeGR4J(
  InputsModel = inputsModel,
  RunOptions = runOptions_val,
  Param = param_vector
)

# Get validation observations
obs_monthly_val <- q %>%
  filter(date >= cal_val_per$valid_start & date <= cal_val_per$valid_end)

# Plot validation results
plot_monthly_comparison(
  best_run_val,
  obs_monthly_val,
  basin_obs_ts$date[indRun_val],
  "Validation Period"
)
```

## Step 8: Analyze Parameter Uncertainty

```{r}
# Plot parameter uncertainty bands
plot_parameter_uncertainty(
  calibration_results,
  n_best = 100,                 # Use top 10 parameter sets
  inputsModel = inputsModel,
  runOptions = runOptions_cal,
  obs_monthly = obs_monthly_cal,
  model_type = "GR4J"
)
```

# 5. RUN MODEL from 2000 - 2024

```{r}

```






