---
title: "Scratch"
author: "Tobias Siegfried, hydrosolutions GmbH"
format: html
editor: visual
---

# Background

Scratch for developing cool code with ariGR hydro modeling

# Change Log

-   2025-04-29: First version

# Libraries

```{r}
library(pacman)
p_Load(tidyverse, airGR)
```


# 5. RUN

## 5.1 Run Model



# X. EXPERIMENTS

There are a few experiments that we can conduct to further improve the model performance. For example, in the computation of the fraction of solid precipitation on the layer, we can calibrate the USACE_Tmin and USACE_Tmax parameters, i.e. the range of the transition temperatures from solid to liquid precipitation.

## X.0 Compute Glazirin Threshold Temperatures for Elevation Bands

This was provided by Andrey. The resulting temperatures are threshold temperatures. > liquid and < solid.
```{r}
z_km <- Basin_Info$ZLayers / 1000

deltaT <- 10

Basin_Info$ThresholdTemp <- (1.25 + 0.016 * z_km + 0.207 * z_km^2) - deltaT
#Basin_Info$ThresholdTemp

# plot threshold temp versus ZLayers in a nice ggplot. For this, we cast ZLayers and Threshold Temp into a data frame
df <- data.frame(ZLayers = Basin_Info$ZLayers, ThresholdTemp = Basin_Info$ThresholdTemp)

ggplot(data = df, aes(x = ZLayers, y = ThresholdTemp)) +
  geom_point() +
  geom_line() +
  labs(x = "Elevation [masl]", y = "Threshold Temperature [°C]") +
  ggtitle("Threshold Temperature vs Elevation, Gleb Glazirin Method") +
  theme_minimal()
```


## X.1 Calibration of sold/liquid preciptation parameters of USACE method

According to the airGR documentation, the USACE method is used to compute the fraction of solid precipitation on the layer. The method requires two parameters: USACE_Tmin and USACE_Tmax. These parameters are used to compute the fraction of solid precipitation on the layer. The source is: USACE (1956), Snow Hydrology, pp. 437. U.S. Army Corps of Engineers (USACE) North Pacific Division, Portland, Oregon, USA.

We use this approach to perturbe elevation specific values. We generate realizations of these parameters using Latin-Hypercube Sampling and then subtract or add these sampled values to the threshold temperatures to arrive at the USACE_Tmin and USACE_Tmax thresholds.

Let us conduct an experiment to calibrate these parameters. 

### Latin-Hypercube Sampling
```{r}
# First install and load the package if you haven't already
# install.packages("lhs")
p_load(lhs)

# Set the number of samples you want
n_samples <- 400  # adjust this number according to needs

# Generate normalized LHS (values between 0 and 1)
set.seed(111)  # for reproducibility
lhs_design <- randomLHS(n = n_samples, k = 2)

# Scale the values to your ranges:
# tas_min variations from Glazirin
Tmin_min <- -5 # -10
T_min_max <- 3 # 2
# USACE_Tmax: from -1 to 5
Tmax_min <- -2 # -1
T_max_max <- 5 # 5


# generate samples
samples <- data.frame(
  USACE_Tmin = lhs_design[,1] * (T_min_max - (Tmin_min)) + (Tmin_min),  # scale to [-10, 2]
  USACE_Tmax = lhs_design[,2] * (T_max_max - (Tmax_min)) + (Tmax_min)     # scale to [-1, 5]
)

# ensure that Tmax is > Tmin
samples <- samples |> 
  filter(USACE_Tmax > USACE_Tmin)

# View first few samples
head(samples)

# You can plot to visualize the distribution
plot(samples$USACE_Tmin, samples$USACE_Tmax,
     xlab = "USACE_Tmin", ylab = "USACE_Tmax",
     main = "Latin Hypercube Samples")
```


### Run Model for Each Sample

Note, this version of the code below is not parallelzied.

```{r}
#| warning: false

# run preparation
# model to use
fun_model <- airGR::RunModel_CemaNeigeGR4J
performance_crit <- list()

# loop over the samples

tictoc::tic()
for (idx in seq(1,nrow(samples))) {
  # print run number
  print(paste0("Run number: ", idx))
  # compute the fraction of solid precipitation on the layer
  solid_frac_pr <- solid_fraction_elevation_layer(Basin_Info, #dates,
                                                  tas = inputsModel_ours$LayerTempMean, 
                                                  tas_min = samples$USACE_Tmin[idx], 
                                                  tas_max = samples$USACE_Tmax[idx]) 
  
  inputsModel_ours$LayerFracSolidPrecip <- solid_frac_pr
  # overwrite original
  inputsModel <- inputsModel_ours
  
  # What we are doing here is to overwrite the setup prepared in the airGR::CreateInputsModel() script. In otherw rods, we just use the data from the data_preparation script.
  
  runOptions_cal <- airGR::CreateRunOptions(FUN_MOD          = fun_model,
                                            InputsModel      = inputsModel,
                                            IndPeriod_Run    = indRun_cal,
                                            IniStates        = NULL,
                                            IniResLevels     = NULL,
                                            IndPeriod_WarmUp = warmup_period_ind,
                                            IsHyst = FALSE,
                                            warnings = FALSE)
  
  inputsCrit_cal <- airGR::CreateInputsCrit(FUN_CRIT    = airGR::ErrorCrit_NSE,
                                            InputsModel = inputsModel,
                                            RunOptions  = runOptions_cal,
                                            transfo = "boxcox",
                                            Obs         = basin_obs_ts$Q_mm_day_no_glacier[indRun_cal],
                                            warnings = FALSE)
  
  calibOptions <- airGR::CreateCalibOptions(FUN_MOD = fun_model, 
                                            FUN_CALIB = airGR::Calibration_Michel)
  
  # calibrate and keep NSE result
  outputsCalib <- airGR::Calibration_Michel(InputsModel = inputsModel,
                                   RunOptions = runOptions_cal,
                                   InputsCrit = inputsCrit_cal,
                                   CalibOptions = calibOptions,
                                   FUN_MOD = fun_model)
  
  performance_crit[[idx]] <- outputsCalib$CritFinal
}
tic_toc <- tictoc::toc()

```

### Parallel Run
```{r}
# Load required packages
p_load(foreach)
p_load(doParallel)

# Set up parallel backend
# Detect number of CPU cores and leave one free for system processes
n_cores <- parallel::detectCores() - 1
# Create and register parallel cluster
cl <- makeCluster(n_cores)
registerDoParallel(cl)

# Start timing
tictoc::tic()

# Parallel execution using foreach
performance_crit <- foreach(idx = 1:nrow(samples),
                          .packages = c("airGR"),  # Add any other required packages
                          .errorhandling = "pass") %dopar% {
  
  # Print run number (note: in parallel execution, prints might be out of order)
  cat(sprintf("Run number: %d\n", idx))
  
  # Compute the fraction of solid precipitation on the layer
  solid_frac_pr <- solid_fraction_elevation_layer(Basin_Info,
                                                  inputsModel_ours$LayerTempMean, 
                                                  tas_min = samples$USACE_Tmin[idx], 
                                                  tas_max = samples$USACE_Tmax[idx]) 
  
  # Create local copy of inputsModel_ours and modify it
  inputsModel <- inputsModel_ours
  inputsModel$LayerFracSolidPrecip <- solid_frac_pr
  
  # Create run options
  runOptions_cal <- airGR::CreateRunOptions(
    FUN_MOD = fun_model,
    InputsModel = inputsModel,
    IndPeriod_Run = indRun_cal,
    IniStates = NULL,
    IniResLevels = NULL,
    IndPeriod_WarmUp = warmup_period_ind,
    IsHyst = FALSE,
    warnings = FALSE
  )
  
  # Create input criteria
  inputsCrit_cal <- airGR::CreateInputsCrit(
    FUN_CRIT = airGR::ErrorCrit_NSE,
    InputsModel = inputsModel,
    RunOptions = runOptions_cal,
    #transfo = "boxcox",
    Obs = basin_obs_ts$Q_mm_day_no_glacier[indRun_cal],
    warnings = FALSE
  )
  
  # Create calibration options
  calibOptions <- airGR::CreateCalibOptions(
    FUN_MOD = fun_model, 
    FUN_CALIB = Calibration_Michel
  )
  
  # Calibrate and return NSE result
  outputsCalib <- Calibration_Michel(
    InputsModel = inputsModel,
    RunOptions = runOptions_cal,
    InputsCrit = inputsCrit_cal,
    CalibOptions = calibOptions,
    FUN_MOD = fun_model
  )
  
  # Return the criterion value
  outputsCalib$CritFinal
}

# Stop timing
time_taken <- tictoc::toc()

# Stop cluster
stopCluster(cl)

# Convert results to a list if they aren't already
performance_crit <- as.list(performance_crit)
```

### Visualize Sampling Results
```{r}
#| warning: false

performance_crit_tbl <- tibble(Tmin = samples$USACE_Tmin,
                               Tmax = samples$USACE_Tmax, 
                               NSE = unlist(performance_crit))

# Create the ggplot first
p <- performance_crit_tbl |>
  ggplot(aes(x = Tmin, y = Tmax, color = NSE)) +
  geom_point(size = 3) +
  scale_color_viridis_c() +  # Note: changed from scale_fill_viridis()
  labs(x = "USACE_Tmin", y = "USACE_Tmax", color = "NSE") +
  theme_minimal()

# Convert to interactive plot
plotly::ggplotly(p)
```

### Calibrate Model with Best Snow / Rain Transition Dataset
```{r}
# Get the best performing parameter set
best_params <- performance_crit_tbl |> 
  filter(NSE == max(NSE)) |> 
  dplyr::select(Tmin, Tmax)

# recompute the fraction of solid precipitation on the layer
solid_frac_pr <- solid_fraction_elevation_layer(Basin_Info, 
                                                  inputsModel_ours$LayerTempMean, 
                                                  tas_min = best_params$Tmin, 
                                                  tas_max = best_params$Tmax) 

inputsModel_ours$LayerFracSolidPrecip <- solid_frac_pr

# overwrite original
inputsModel <- inputsModel_ours

runOptions_cal <- airGR::CreateRunOptions(FUN_MOD          = fun_model,
                                          InputsModel      = inputsModel,
                                          IndPeriod_Run    = indRun_cal,
                                          IniStates        = NULL,
                                          IniResLevels     = NULL,
                                          IndPeriod_WarmUp = warmup_period_ind,
                                          IsHyst           = FALSE,
                                          warnings         = FALSE)

inputsCrit_cal <- airGR::CreateInputsCrit(FUN_CRIT         = fun_crit,
                                          InputsModel      = inputsModel,
                                          RunOptions       = runOptions_cal,
                                          transfo          = transfo,
                                          Obs              = q_obs_cal,
                                          warnings         = FALSE)

calibOptions <- airGR::CreateCalibOptions(FUN_MOD = fun_model, 
                                          FUN_CALIB = Calibration_Michel)

# calibrate and keep NSE result
outputsCalib <- Calibration_Michel(InputsModel = inputsModel,
                                   RunOptions = runOptions_cal,
                                   InputsCrit = inputsCrit_cal,
                                   CalibOptions = calibOptions,
                                   FUN_MOD = fun_model)

param <- outputsCalib$ParamFinalR

# switch call of function depending on fun_model
if (model_2_use == "GR4J") {
  runResults_cal <- RunModel_CemaNeigeGR4J(InputsModel = inputsModel,
                                           RunOptions = runOptions_cal,
                                           Param = param)
} else if (model_2_use == "GR5J") {
  runResults_cal <- RunModel_CemaNeigeGR5J(InputsModel = inputsModel,
                                           RunOptions = runOptions_cal,
                                           Param = param)
} else if (model_2_use == "GR6J") {
  runResults_cal <- RunModel_CemaNeigeGR6J(InputsModel = inputsModel,
                                           RunOptions = runOptions_cal,
                                           Param = param)
}
#runResults_cal <- RunModel_CemaNeigeGR6J(InputsModel = inputsModel,
#                                         RunOptions = runOptions_cal,
#                                         Param = param)

plot(runResults_cal, Qobs = basin_obs_ts$Q_mm_day[indRun_cal])

OutputsCrit <- ErrorCrit_NSE(InputsCrit = inputsCrit_cal, OutputsModel = runResults_cal)
OutputsCrit
```

### Validation Run
```{r}
runOptions_val <- airGR::CreateRunOptions(FUN_MOD          = fun_model,
                                          InputsModel      = inputsModel,
                                          IndPeriod_Run    = indRun_val,
                                          IniStates        = runResults_cal$StateEnd,
                                          IniResLevels     = NULL)

inputsCrit_val <- airGR::CreateInputsCrit(FUN_CRIT         = fun_crit,
                                          InputsModel      = inputsModel,
                                          RunOptions       = runOptions_val,
                                          transfo          = transfo,
                                          Obs              = q_obs_val,
                                          Weights          = weights)

# switch call of function depending on fun_model
if (model_2_use == "GR4J") {
  runResults_val <- RunModel_CemaNeigeGR4J(InputsModel     = inputsModel,
                                           RunOptions      = runOptions_val,
                                           Param           = param)
} else if (model_2_use == "GR5J") {
  runResults_val <- RunModel_CemaNeigeGR5J(InputsModel     = inputsModel,
                                           RunOptions      = runOptions_val,
                                           Param           = param)
} else if (model_2_use == "GR6J") {
  runResults_val <- RunModel_CemaNeigeGR6J(InputsModel     = inputsModel,
                                           RunOptions      = runOptions_val,
                                           Param           = param)
}

plot(runResults_val, Qobs = basin_obs_ts$Q_mm_day[indRun_val])

OutputsCrit <- ErrorCrit_NSE(InputsCrit = inputsCrit_val, OutputsModel = runResults_val)
OutputsCrit
                                          
```


